Classification of RDF Constraint Types Ensuring High Quality of Metadata and Data

### 2015-04-27 paper/paper7/paper7.tex
motivation

By its nature, person-level data is highly confidential and access is often only permitted for qualified researchers who must apply for access. 
Researchers typically represent their results as aggregated data in form of multi-dimensional tables with only a few columns; so-called \emph{variables} such as \emph{sex} or \emph{age}.
%In performing their research, the detailed person-level
%data is aggregated into multi-dimensional tables which answer particular research questions.
Aggregated data, which answers particular research questions, is derived from person-level data by statistics on groups or aggregates such as frequencies and arithmetic means.
%The SDMX metadata standard – used as the basis for \emph{QB} – and DDI have traditionally made efforts to align their content. 
%Similarly, some of the developers of \emph{Disco} were also involved in the development of \emph{QB}, 
%allowing the RDF versions of these standards to retain that alignment.
%While \emph{Disco} and \emph{QB} provide terms for the description of data sets, 
%both on a different level of aggregation, 
%the \emph{Data Catalog Vocabulary (DCAT)}\footnote{\url{http://www.w3.org/TR/vocab-dcat/}} enables the representation of these data sets inside of data collections like repositories, catalogs, or archives. 
%The relationship between data collections and their contained data sets is useful, since such collections are a typical entry point when searching for data.
%Although, in most cases aggregated data is still published in form of PDFs, 
%it is more and more common to publish aggregated data as CSV files,
The purpose of publicly available aggregated data is to get a first overview and to gain an interest in further analyses on the underlying person-level data.
Aggregated data is more and more published in form of CSV files,
allowing to perform data calculations.
%For these calculations, definitions of the columns are needed (e.g. is a given variable interpreted numerically or as a string).
%\emph{Physical Data Description (PHDD)}\footnote{\url{https://github.com/linked-statistics/physical-data-description}} is a vocabulary to represent data in tabular format in RDF.\tb{tabular data} 
%The data could be either represented in records with character-separated values (CSV) or in records with fixed length.

For more detailed analyses, researchers refer to person-level data including additional variables needed to answer subsequent research questions
%Although not that common, the other direction is also possible, 
%i.e. researchers may use metadata on person-level data to search for aggregated data.
like the comparison of studies between countries.
A \emph{study} represents the process by which a data set was generated or collected.
Eurostat\footnote{\url{http://ec.europa.eu/eurostat}}, the statistical office of the European Union, provides research findings in form of aggregated data (downloadable as CSV files) and its metadata at European level that enable comparisons between countries.
%This way, researchers get promising findings (in form of published tables with a few columns).

%We use the availability of childcare services in European Union member states by year, duration, and child age leading to subsequent research questions as running example of this paper.
%The first step is to determine in which countries the unemployment rate is collected and which other variables of each country-specific study are theoretically comparable and can therefore be used to answer the underlying research question.
The variable \emph{formal childcare}\footnote{Aggregated data and its metadata is available at: \url{http://ec.europa.eu/eurostat/web/products-datasets/-/ilc_caindformal}} %(in contrast to childcare at home)
captures the measured availability of childcare services in percent over the population in European Union member states by 
the variables \emph{year}, \emph{duration} (in hours per week), \emph{age} of the child, and \emph{country}.
Variables are constructed out of values (of one or multiple datatypes) and/or code lists.
The variable \emph{age}, e.g., may be represented by values of the datatype \emph{xsd:nonNegativeInteger}, or by a code list including multiple age clusters (such as '0 to 10' and '11 to 20').

To determine if variables measuring \emph{age} 
- collected for different countries (\emph{$age_{DE}$}, \emph{$age_{UK}$}) - 
are comparable, both content-driven and technology-driven constraints are validated.
Content-driven constraints ensure that the data is consistent with the intended syntax, semantics, and integrity of vocabularies' data models and 
technology-driven constraints can be generated completely automatically out of vocabularies' data models. 
Examples for content-driven constraints are to investigate (1) if variables are represented in a compatible way,
i.e. are the variables' code lists theoretically comparable, and (2) if variables' code lists are properly structured.
With technology-driven constraints, it can be validated (1) if variable definitions are available and (2) if for each code an associated category (a human-readable label) is specified.

-----

Portals harvest metadata (as well as publicly available data) from multiple RDF data providers.
To ensure high quality, (meta)data must satisfy certain criteria - specified in terms of RDF constraints.  
%After validating the metadata according to these constraints, portals offer added values to their customers, e.g., by searching over and comparing metadata of multiple providers.

@@

%\textbf{Organizations, Hierarchies, and Classifications.}
%\textbf{Reusing SKOS.}
The \emph{Simple Knowledge Organization System (SKOS)} is reused multiple times to build \emph{SBE} vocabularies.
Variables are constructed out of values and/or (un)ordered code lists.
The codes of the variable \emph{Education at pre-school} (number of education hours per week) are modeled as \emph{skos:Concepts} and 
a \emph{skos:OrderedCollection} organizes them in a particular order within a \emph{skos:memberList}.
A variable may be associated with a theoretical concept (\emph{skos:Concept}). 
\emph{skos:narrower} builds the hierarchy of theoretical concepts within a \emph{skos:ConceptScheme} of a series.
The variable \emph{Education at pre-school} is assigned to the theoretical concept \emph{Child Care} which is the narrower concept of \emph{Education} - one of the top concepts of the series \emph{EU-SILC}.
Controlled vocabularies (\emph{skos:ConceptScheme}), serving as extension and reuse mechanism,
organize types (\emph{skos:Concept}) of descriptive statistics (\emph{disco:SummaryStatistics}) like minimum, maximum, and arithmetic mean.
%\emph{XKOS}\footnote{\url{https://github.com/linked-statistics/xkos}} is a SKOS extension to describe formal statistical classifications like the International Standard Classification of Occupations (\emph{ISCO}). 
%and the Statistical Classification of Economic Activities in the European Community \emph{NACE}.

@@

A very common research question is to compare variables of multiple studies or countries (constraint type: \emph{comparison}).
To compare variables 
(1) their code lists must be structured properly and
(2) their code lists must either be identical or at least similar.
If a researcher only wants to get a first overview over comparable variables (use case 1), 
covering the first constraint may be sufficient.
Thus, the severity level of the first constraint is stronger ($\mathcal{SL}_{2}$) than the one for the second constraint ($\mathcal{SL}_{0}$).
If the intention of the researcher is to perform more detailed comparisons (use case 2), however, the violation of the second constraint is getting more serious
and the user may raise its severity level.
%\textbf{Comparison.}
%A very common research question is to compare variables of multiple studies or countries (constraint type: \emph{comparison}).
%To compare variables, 
%(1) variables and (2) variable definitions must be present,
%(3) code lists must be structured properly,
%(4) for each code an associated category must be specified, and
%(5) code lists must either be identical or at least similar.
%If a researcher wants to get a first overview over comparable variables (use case 1), 
%covering the first three constraints may be sufficient.
%Thus, the severity level of the first three constraints is stronger ($\mathcal{SL}_{2}$) than for the last two constraints ($\mathcal{SL}_{1}$ and $\mathcal{SL}_{0}$).
%If the intention of the researcher is to perform more detailed comparisons (use case 2), however, the violation of the remaining two constraints is getting more serious.

>>>>

By its nature, person-level data is highly confidential and access is often only permitted for qualified researchers who must apply for access. 
Researchers typically represent their results as aggregated data in form of multi-dimensional tables with only a few columns; so-called \emph{variables} such as \emph{sex} or \emph{age}.
%In performing their research, the detailed person-level
%data is aggregated into multi-dimensional tables which answer particular research questions.
Aggregated data, which answers particular research questions, is derived from person-level data by statistics on groups or aggregates such as frequencies and arithmetic means.
%The SDMX metadata standard – used as the basis for \emph{QB} – and DDI have traditionally made efforts to align their content. 
%Similarly, some of the developers of \emph{Disco} were also involved in the development of \emph{QB}, 
%allowing the RDF versions of these standards to retain that alignment.
%While \emph{Disco} and \emph{QB} provide terms for the description of data sets, 
%both on a different level of aggregation, 
%the \emph{Data Catalog Vocabulary (DCAT)}\footnote{\url{http://www.w3.org/TR/vocab-dcat/}} enables the representation of these data sets inside of data collections like repositories, catalogs, or archives. 
%The relationship between data collections and their contained data sets is useful, since such collections are a typical entry point when searching for data.
%Although, in most cases aggregated data is still published in form of PDFs, 
%it is more and more common to publish aggregated data as CSV files,
The purpose of publicly available aggregated data is to get a first overview and to gain an interest in further analyses on the underlying person-level data.
Aggregated data is more and more published in form of CSV files,
allowing to perform data calculations.
%For these calculations, definitions of the columns are needed (e.g. is a given variable interpreted numerically or as a string).
%\emph{Physical Data Description (PHDD)}\footnote{\url{https://github.com/linked-statistics/physical-data-description}} is a vocabulary to represent data in tabular format in RDF.\tb{tabular data} 
%The data could be either represented in records with character-separated values (CSV) or in records with fixed length. 
Portals harvest metadata (as well as publicly available data) from multiple RDF data providers.
To ensure high quality, (meta)data must satisfy certain criteria - specified in terms of RDF constraints.

For more detailed analyses, researchers refer to person-level data including additional variables needed to answer subsequent research questions
%Although not that common, the other direction is also possible, 
%i.e. researchers may use metadata on person-level data to search for aggregated data.
like the comparison of studies between countries.
A \emph{study} represents the process by which a data set was generated or collected.
Eurostat\footnote{\url{http://ec.europa.eu/eurostat}}, the statistical office of the European Union, provides research findings in form of aggregated data (downloadable as CSV files) and its metadata at European level that enable comparisons between countries.
%This way, researchers get promising findings (in form of published tables with a few columns).
%We use the availability of childcare services in European Union member states by year, duration, and child age leading to subsequent research questions as running example of this paper.
%The first step is to determine in which countries the unemployment rate is collected and which other variables of each country-specific study are theoretically comparable and can therefore be used to answer the underlying research question.
The variable \emph{formal childcare}\footnote{Aggregated data and its metadata is available at: \url{http://ec.europa.eu/eurostat/web/products-datasets/-/ilc_caindformal}} %(in contrast to childcare at home)
captures the measured availability of childcare services in percent over the population in European Union member states by 
the variables \emph{year}, \emph{duration} (in hours per week), \emph{age} of the child, and \emph{country}.
Variables are constructed out of values (of one or multiple datatypes) and/or code lists.
The variable \emph{age}, e.g., may be represented by values of the datatype \emph{xsd:nonNegativeInteger} or by a code list of age clusters (e.g., '0 to 10' and '11 to 20').

To determine if variables measuring \emph{age} 
- collected for different countries (\emph{$age_{DE}$}, \emph{$age_{UK}$}) - 
are comparable, diverse constraints are checked:
(1) variable definitions must be available, 
(2) for each code a human-readable label has to be specified,
(3) code lists must be structured properly, and
(4) code lists must either be identical or at least similar.
If a researcher only wants to get a first overview over comparable variables (use case 1), 
covering the first three constraints may be sufficient,
i.e., the violation of the first three constraints is more serious than the violation of the last constraint.
If the intention of the researcher is to perform more sophisticated comparisons (use case 2), however, the user may raise the severity level of last constraint.

%After validating the metadata according to these constraints, portals offer added values to their customers, e.g., by searching over and comparing metadata of multiple providers.

@@

%\textbf{Organizations, Hierarchies, and Classifications.}
%\textbf{Reusing SKOS.}
The \emph{Simple Knowledge Organization System (SKOS)} is reused multiple times to build \emph{SBE} vocabularies.
%Variables are constructed out of values and/or (un)ordered code lists.
The codes of the variable \emph{Education at pre-school} (number of education hours per week) are modeled as \emph{skos:Concepts} and 
a \emph{skos:OrderedCollection} organizes them in a particular order within a \emph{skos:memberList}.
A variable may be associated with a theoretical concept (\emph{skos:Concept}). 
\emph{skos:narrower} builds the hierarchy of theoretical concepts within a \emph{skos:ConceptScheme} of a series.
The variable \emph{Education at pre-school} is assigned to the theoretical concept \emph{Child Care} which is the narrower concept of \emph{Education} - one of the top concepts of the series \emph{EU-SILC}.
Controlled vocabularies (\emph{skos:ConceptScheme}), serving as extension and reuse mechanism,
organize types (\emph{skos:Concept}) of descriptive statistics (\emph{disco:SummaryStatistics}) like minimum, maximum, and arithmetic mean.
%\emph{XKOS}\footnote{\url{https://github.com/linked-statistics/xkos}} is a SKOS extension to describe formal statistical classifications like the International Standard Classification of Occupations (\emph{ISCO}). 
%and the Statistical Classification of Economic Activities in the European Community \emph{NACE}.

@@

%A very common research question is to compare variables of multiple studies or countries (constraint type: \emph{comparison}).
%To compare variables 
%(1) their code lists must be structured properly and
%(2) their code lists must either be identical or at least similar.
%If a researcher only wants to get a first overview over comparable variables (use case 1), 
%covering the first constraint may be sufficient.
%Thus, the severity level of the first constraint is stronger ($\mathcal{SL}_{2}$) than the one for the second constraint ($\mathcal{SL}_{0}$).
%If the intention of the researcher is to perform more detailed comparisons (use case 2), however, the violation of the second constraint is getting more serious
%and the user may raise its severity level.
%\textbf{Comparison.}
%A very common research question is to compare variables of multiple studies or countries (constraint type: \emph{comparison}).
%To compare variables, 
%(1) variables and (2) variable definitions must be present,
%(3) code lists must be structured properly,
%(4) for each code an associated category must be specified, and
%(5) code lists must either be identical or at least similar.
%If a researcher wants to get a first overview over comparable variables (use case 1), 
%covering the first three constraints may be sufficient.
%Thus, the severity level of the first three constraints is stronger ($\mathcal{SL}_{2}$) than for the last two constraints ($\mathcal{SL}_{1}$ and $\mathcal{SL}_{0}$).
%If the intention of the researcher is to perform more detailed comparisons (use case 2), however, the violation of the remaining two constraints is getting more serious.
